Conflict-Free Vertex Coloring Of Planar Graphs

### 2017-03-21 paper/paper.tex
Adding more work on CF coloring

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example.pdf}
	\caption{A simple, undirected graph $H$}\label{fig:graph}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example-vcp.pdf}
	\caption{A minimum vertex coloring of $H$}\label{fig:vcp-example}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example-cfcp.pdf}
	\caption{A minimum conflict-free coloring of $H$}\label{fig:cfcp-example}
\end{figure}

@@

An example of a conflict-free coloring is shown in figure \ref{fig:cfcp-example}. It is worth noting that all proper vertex colorings are also conflict-free colorings. This is because in a proper vertex coloring, every vertex is its own conflict-free neighbor \cite{abel2017three}. We observe from figures \ref{fig:vcp-example} and \ref{fig:cfcp-example}, $\chi(H) = 3$ and $\chi_{CF}(H) = 2$.

The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $H$ where $\gamma(H) = 2$.

@@

As shown with graph $H$ in section \ref{sec:background}, even though all proper vertex colorings are also conflict-free colorings, fewer colors can usually be used. This leads us to use heuristics that differ from the vertex coloring problem. Abel et al. \cite{abel2017three} present an efficient heuristic to color certain general graphs with $k$ colors in a conflict-free manner. They call this heuristic \emph{iterated elimination of distance-3-sets}.

\subsection{Guaranteeing CF k-Colorability}
It is wasteful to spend time coloring a graph that cannot be conflict-free colored with $k$ colors. This leads Abel et al. \cite{abel2017three} to provide sufficient criterion to guarantee the conflict-free k-colorability of a graph. To introduce this criterion, we need to define a few more terms.

@@

\section{Acknowledgments}
\cite{abel2017three}

>>>>

\newtheorem{theorem}{Theorem}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example.pdf}
	\caption{A simple, undirected graph $G_1$}\label{fig:graph}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example-vcp.pdf}
	\caption{A minimum vertex coloring of $G_1$}\label{fig:vcp-example}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=7cm]{../figures/example-cfcp.pdf}
	\caption{A minimum conflict-free coloring of $G_1$}\label{fig:cfcp-example}
\end{figure}

@@

An example of a conflict-free coloring is shown in figure \ref{fig:cfcp-example}. It is worth noting that all proper vertex colorings are also conflict-free colorings. This is because in a proper vertex coloring, every vertex is its own conflict-free neighbor \cite{abel2017three}. We observe from figures \ref{fig:vcp-example} and \ref{fig:cfcp-example}, $\chi(G_1) = 3$ and $\chi_{CF}(G_1) = 2$.

The domination number of $G$, denoted $\gamma(G)$, is the size of a minimum dominating set of G. A dominating set is a subset $D$ of $V$ such that every vertex not in $D$ is adjacent to at least one vertex in $D$. The \emph{conflict-free domination number} for some $k$, denoted $\gamma_{CF}^k(G)$, is the minimum number of vertices that have to be colored in a conflict-free k-coloring of $G$. An example of a dominating set is $\{3, 5\}$ of graph $G_1$ where $\gamma(G_1) = 2$.

@@

As shown with graph $G_1$ in section \ref{sec:background}, even though all proper vertex colorings are also conflict-free colorings, fewer colors can usually be used. This leads us to use heuristics that differ from the vertex coloring problem. Abel et al. \cite{abel2017three} present an efficient heuristic to color certain general graphs with $k$ colors in a conflict-free manner. They call this heuristic \emph{iterated elimination of distance-3-sets}.

\subsection{Guaranteeing CF k-Colorability}
It is wasteful to spend time coloring a graph that cannot be conflict-free colored. This leads Abel et al. to provide sufficient criterion to guarantee the conflict-free k-colorability of a certain graphs. To introduce this criterion, we need to introduce a few specific graphs. A complete graph is a simple, undirected graph where every pair of distinct vertices is connected by an edge. A complete graph on $n$ vertices is denoted as $K_n$. The graph $K_n^{-3}$ is the graph obtained by removing any three edges forming a single triangle, i.e. \(K_n^{-3} =  \left ( V(K_n), E(K_n) \setminus E(K_3) \right ) \). A graph $H$ is called a \emph{minor} of a graph $G$ if $H$ can be formed from $G$ by deleting edges and vertices and by contracting edges. \cite{abel2017three,bondy1976graph}

\begin{theorem} \label{thm:criterion}
Let $G$ be a graph and $k \geq 1$. If $G$ has neither $K_{k+2}$ nor $K_{k+3}^{-3}$ as a minor, $G$ has a conflict-free coloring that can be found in polynomial time using iterated elimination of distance-3-sets.
\end{theorem}

Abel et al. give a criterion that guarantees k-colorability which is stated in theorem \ref{thm:criterion}. In the proof of this theorem, found in \cite{abel2017three}, the authors prove that their heuristic will always generate a k-colorable graph when the specified criterion is met. To illustrate this heuristic, we will generate a graph that fits the specified criterion and apply the iterated elimination of distance-3-sets algorithm defined as algorithm \ref{alg:elimination}.

% %%%%% REMOVE
% \newpage
% %%%%%

\subsection{Setting Up Example CF Coloring}
We will demonstrate meeting the criterion to utilize this algorithm using the graph $G_2$ in figure \ref{fig:criterion}.

\begin{figure}[h]
	\centering
	\includegraphics[width=6cm]{../figures/criterion.pdf}
	\caption{A simple graph $G_2$}\label{fig:criterion}
\end{figure}

We will be focusing on the simplest example, $k = 1$. Graph $G_2$ is a simple graph with 6 vertices. It does not, however, fit our criterion; it actually breaks both properties as it has both $K_{k+2}$ and $K_{k+4}^{-3}$ as a minor. For $k = 1$, this means our graph cannot contain $K_{3}$ or $K_{4}^{-3}$ as a minor. These graphs are shown in figure \ref{fig:bad-criterion}.

\begin{figure}[h]
	\centering
	\includegraphics[width=6cm]{../figures/bad-criterion.pdf}
	\caption{Graphs $K_{3}$ and $K_{4}^{-3}$, respectively}\label{fig:bad-criterion}
\end{figure}

We can see that $G_2$ has both of these as minors. The transformation needed to see the minors is shown in figure \ref{fig:criterion-minor}. By contracting the dotted edges, vertices 1 and 4 would combine as well as vertices 3 and 5. We would have both $K_{3}$ and $K_{4}^{-3}$. Thus, this graph cannot be conflict-free colored with one color.

\begin{figure}[h]
	\centering
	\includegraphics[width=6cm]{../figures/criterion-minor.pdf}
	\caption{How to transform $G_2$ to see the minors}\label{fig:criterion-minor}
\end{figure}

@@

\section{Acknowledgments}
Thanks to Peter Dolan and Elena Machkasova for their feedback on this paper.
